---
title: "Run Targets - lakeSR pipeline"
author: "ROSSyndicate"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# Purpose

This script loads all necessary packages to run the {targets} pipeline for 
lakeSR then runs it! 

Important note: before rendering this document or running this workflow, you 
will need to authenticate your Earth Engine credentials in the *zsh* terminal. 
See the README in this repository for details on this action.

## Install necessary packages

Define package installer function

```{r package_installer}
package_installer <- function(x) {
  if (x %in% installed.packages()) {
    print(paste0("{", x ,"} package is already installed."))
    } else {
      install.packages(x)
      print(paste0("{", x ,"} package has been installed."))
    }
  }
```

List packages that need to be checked for install, and walk the function along them all.

```{r walk_package_installer}
packages <- c("tidyverse",
              "targets",
              "tarchetypes",
              "nhdplusTools",
              "polylabelr",
              "sf",
              "tigris",
              "reticulate")
lapply(packages, package_installer)
```

## Authenticate Earth Engine

This workflow requires authentication of your earth engine account. There are 
additional dependencies listed in the README file as well as links for 
troubleshooting. 

## Run the targets pipeline and output a network graph.

```{r run_targets}
library(targets)

# run the pipeline
tar_make()

```

### Create a network diagram of the workflow.

```{r see_targets_net}
tar_visnetwork()
```

