---
editor_options:
  markdown:
    wrap: 80
bibliography: references.bib
params:
  cfg: NA
  LS_files: NA
---

```{r chunk-opts-6, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      error = FALSE)

library(tidyverse)
library(googledrive)
library(feather)

drive_auth(params$cfg$google_email)

#LS_files <- read_csv("c_collate_Landsat_data/out/raw_collated_files_drive_ids_2024-12-16.csv")
```

# Calculating Intermission Handoffs {#handoff}

It is necessary to standardize band values between Landsat missions in order to
create a robust timeseries. @maciel2023 notes the issues in using the Landsat
Surface Reflectance product as a timeseries without careful handling of data
between missions. This standardization process attempts to address changes in
sensor spectral response and atmospheric correction procedures. For the purposes
of AquaSat v2, we call this standardization process "intermission handoffs".
Intermission handoffs are defined by the data obtained in the overlapping period
of time between two adjacent-in-time missions (details provided in Section
\@ref(define-handoff). In AquaSat v2, we standardize values relative to Landsat
7 (`corr7`) and Landsat 8 (`corr8`). The Landsat 7 intermission handoffs can be
applied the entire Landsat 4-9 because it overlapped with early Landsat (4/5)
and modern Landsat (8/9). For the `corr7` standardization, we treat Landsat 4
and 5 as interchangable, given that the satellites contained the same payload
and radiometric resolution. We also treat Landsat 8 and 9 as interchangable
despite the slight differences in spectral response and radiometric resolution.
Landsat 8 handoffs (`corr8`) can only be applied to Landsat 7 and 9 due to
mission overlap, so the standardized data timeseries is shorter; however, it may
be more robust for analyses focused in the recent past that rely on Landsat 9
data given the previously-mentioned caveats for application of `corr7` to that
mission.

## Raw lakeSR stacks {#raw-lakesr}

The data exported from the GEE tasks described in Section \@ref(LS-SRST-pull)
were collated into large .feather files per DSWE type (DSWE1 and DSWE1a) and per
Landsat mission. These files are stored in Google Drive for posterity and to add
functionality to this workflow. The Google Drive ids can be found at the file
path `c_collate_Landsat_data/out/` and can be accessed without special
permissions by downstream users. Many of these files are quite large (on the
order of many GBs) and are very difficult to handle outside of a programming or
database environment. We provide these files for advanced users who wish to make
changes to the QAQC or intermission handoff procedures.

## Post-hoc filters {#post-hoc-filters}

We acknowledge that our workflow may still result in erroneous or misleading SR
values, to address that we have a few post-hoc filters that we apply to the
stack to reduce uncertainty in the remote sensing data as much as possible.

pCount_dswe1 \>= 8 - change from ASv1, with 10. Due to the more rigorous masks,
we think this is acceptable.

all med_band data must be \< 0.2, which aligns with the expected optical
properties of water and helps reduce errors associated with sun glint.

either NIR or SWIR1 and SWIR2 must be \< 0.1, this is to remove any extracted
samples where sun glint has likely affected the data. In aquatic environments,
NIR, SWIR1, and SWIR2 surface reflectance values should be very low and these
bands can be used for detecting sun glint affected areas [e.g. @mondejar2019,
@vanhellemont2019]. Since NIR bands can be elevated in high-sediment waters
[e.g., @doxaran2002] and SWIR can be elevated in high chlorophyll environments
[e.g, @hu2009], and we did not wish to bias our dataset and remove data of this
nature, we embraced this conditional approach. Very few instances occur where
both NIR and at least one of th SWIR bands were greater than 0.1 after filtering
for all bands \<0.2.

if n_masked \<-0.01 \> pCount_dswe1/5, drop

```{r step_function}
make_LS_QA_step_graphic <- function(mission, 
                                    drive_info, landsat, dswe, version_date) {
  # get specific file
  filtered_files <- drive_info %>%
    filter(str_detect(name, landsat) & 
             str_detect(name, dswe) &
             str_detect(name, version_date))
  
  # make sure there is only one file
  if (nrow(filtered_files) > 1) {
    stop("Landsat, DSWE, and version_date arguments resulted in more than one file. Check format.")
  }
  
  tmp_file <- tempfile()
  drive_download(as_id(filtered_files$id[1]), tmp_file)
  data <- read_feather(tmp_file)
  unlink(tmp_file)
  
  # filter for at least 8 pixels
  valid_px <- data %>% 
    filter(pCount_dswe1 >= 8)
  # filter for all optical bands < 0.2
  bands <- names(valid_px)[grepl("med_", names(valid_px))]
  optical <- bands[!bands %in% "med_SurfaceTemp"]
  valid_band <- valid_px %>% 
    filter(if_all(all_of(optical),
                  ~ . < 0.2))
  # filter thermal for > 273.15 (above freezing)
  gt_zero <- valid_band %>% 
    filter(med_SurfaceTemp > 273.15)
  # filter for nir/swir thresholds
  remove_glint <- gt_zero %>% 
    filter(med_Nir < 0.1 | (med_Swir1 < 0.1 & med_Swir2 < 0.1))
  

  row_summary <- tibble(all_data = nrow(data),
    valid_px = nrow(valid_px),
    valid_band = nrow(valid_band),
    gt_zero = nrow(gt_zero),
    remove_glint = nrow(remove_glint)) %>% 
    pivot_longer(cols = all_data:remove_glint) 
  
  drop_reason <- tibble(all_data = " unfiltered Landsat data",
    valid_px = " at least 8 pixels contribute to summary",
    valid_band = " all optical bands < 0.2",
    gt_zero = " thermal band > 273.15",
    remove_glint = " NIR/SWIR thresholds met") %>% 
    pivot_longer(cols = all_data:remove_glint,
                 values_to = "reason") 
  
  drops <- full_join(row_summary, drop_reason) %>% 
    mutate(name = factor(name, levels = c("remove_glint",
                                          "gt_zero",
                                          "valid_band",
                                          "valid_px",
                                          "all_data")),
           lab = paste0(reason, ": ", format(value, big.mark = ","), " records"))
  
  ggplot(drops) +
    geom_bar(aes(x = name, y = value, fill = name),
             stat = "identity")  +
    geom_text_repel(aes(x = name, y = 0.1, label = lab),
                    bg.color = "white", bg.r = 0.15,
                    point.size = NA,
                    xlim = c(-Inf, Inf),
                    ylim =  c(-Inf, Inf),
                    nudge_y = 100,
                    hjust = "left") +
    labs(title = paste0("Summary of ", mission, " data QA records"), 
         x = NULL, y = NULL) +
    scale_fill_manual(values = viridis(n = nrow(drops),
                                       direction = -1)) +
    scale_x_discrete(drop = F) +
    coord_flip() +
    theme_bw() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          plot.title = element_text(size = 14, face = "bold", hjust = 0.5), 
          legend.position = "none")
  }

```

```{r LS4_step, eval = F}
make_LS_QA_step_graphic(mission = "Landsat 4 - DSWE1 - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LT04", 
                        dswe = "DSWE1_", 
                        version_date = "2024-12-16")
make_LS_QA_step_graphic(mission = "Landsat 4 - DSWE1a - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LT04", 
                        dswe = "DSWE1a_", 
                        version_date = "2024-12-16")
```

```{r LS5_step, eval = F}
make_LS_QA_step_graphic(mission = "Landsat 5 - DSWE1 - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LT05", 
                        dswe = "DSWE1_", 
                        version_date = "2024-12-16")
make_LS_QA_step_graphic(mission = "Landsat 5 - DSWE1a - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LT05", 
                        dswe = "DSWE1a_", 
                        version_date = "2024-12-16")
```

```{r LS7_step, eval = F}
make_LS_QA_step_graphic(mission = "Landsat 7 - DSWE1 - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LE07", 
                        dswe = "DSWE1_", 
                        version_date = "2024-12-16")
make_LS_QA_step_graphic(mission = "Landsat 7 - DSWE1a - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LE07", 
                        dswe = "DSWE1a_", 
                        version_date = "2024-12-16")
```

```{r LS8_step, eval = F}
make_LS_QA_step_graphic(mission = "Landsat 8 - DSWE1 - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LC08", 
                        dswe = "DSWE1_", 
                        version_date = "2024-12-16")
make_LS_QA_step_graphic(mission = "Landsat 8 - DSWE1a - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LC08", 
                        dswe = "DSWE1a_", 
                        version_date = "2024-12-16")
```

```{r LS9_step, eval = F}
make_LS_QA_step_graphic(mission = "Landsat 9 - DSWE1 - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LC09", 
                        dswe = "DSWE1_", 
                        version_date = "2024-12-16")
make_LS_QA_step_graphic(mission = "Landsat 9 - DSWE1a - lakeSR", 
                        drive_info = LS_files, 
                        landsat = "LC09", 
                        dswe = "DSWE1a_", 
                        version_date = "2024-12-16")
```

## Calculating handoffs {#define-handoff}

We follow the practice described in @gardner2021 to calculate intermission
handoffs, where all band values during the period of overlap between two
missions (e.g. 1999-04-15 through 2011-11-18 for Landsat 5 and 7 - see Section
\@ref(valid-dates)) are considered. These data are summarized to each mission's
1st-99th percentile value per band, and the handoff coefficients between
missions are defined by the second-order polynomial relationship between them.
Because we use a second-order polynomial, all values outside of the 1st and 99th
percentile values used to define the intermission handoff are flagged and should
be used with extreme caution. All `corr7` and `corr8` relationships and their
derived coefficients are provided in the sections below.

### `corr7` handoff coefficients

[[pull figs from functions here]]

[[insert table of calcs, \# of points, etc]]

### `corr8` handoff coefficients

[[pull figs from functions here]]

[[insert table of calcs, \# of points, etc]]
