---
editor_options:
  markdown:
    wrap: 80
bibliography: references.bib
params:
  LS5_for57: NA
  LS7_for57: NA
  LS7_for78: NA
  LS8_for78: NA
  LS8_for89: NA
  LS9_for89: NA
  LS57_match: NA
  LS78_match: NA
  LS89_match: NA
tables: TRUE
---

# Calculating intermission handoffs {#define-handoff}

It is necessary to standardize band values between Landsat missions in order to
create a robust timeseries. @maciel2023 notes the issues in using the Landsat
Surface Reflectance product as a timeseries without careful handling of data
between missions. This standardization process attempts to address changes in
sensor spectral response and atmospheric correction procedures. For the purposes
of AquaSat v2, we call this standardization process "intermission handoffs". We
implement two versions of intermission handoffs: the method described in
@roy2016 ("Roy method") and an adapted version of that described in @gardner2021
("Gardner method").

In AquaSat v2, we provide coefficients to standardize remote sensing values
relative to Landsat 7 (`corr7`) and Landsat 8 (`corr8`). The Landsat 7
intermission handoffs create a continuous record of remote sensing from Landsat
4 through 9 relative to Landsat 7 since Landsat 4 and 5 and Landsat 8 and 9 can
be treated as interchangable due to the similarity in payload and radiometric
resolution (see section \@ref(establish) for details and caveats). Landsat 8
handoffs (`corr8`) can only be applied to Landsat 7 and 9 due to mission
overlap, so the standardized data timeseries is shorter.

## Roy method

The Roy method for calculating intermission handoffs uses paired images at
specific sites, where the reflectance data are obtained are from two missions
(e.g. Landsat 7 and Landsat 8) separated by one day at a specific location
within the overlapping paths in the WRS2 path-row framework (see figure
@ref(path-overlap-roy)).

![An example of two overlapping WRS2 paths (grey area) separated by one day. Any
location with data for both images would be considered a paired data point to
define the Roy method intermission
handoffs.](images/path_overlap_roy.png){alt="An example of two overlapping WRS2 paths (grey area) separated by one day. Any location with data for both images would be considered a paired data point to define the Roy method intermission handoffs."}

While @roy2016 implemented this method on a pixel-by-pixel basis, we implement
using the median band value per site. This is used in place of the explicit
filters described in the Roy method (saturated pixels, cloudy/snowy pixels, and
pixels with value changes greater than the changes in atmospheric correction),
as we have implemented masks and QA filters to reduce these sources of error.
Handoff coefficients are defined by the ordinary least squares regression line
of the Landsat 5 against the Landsat 7 data and Landsat 7 against the Landsat 8
data (see Table \@ref(tab:roy-table)). For the purposes of this documentation,
we only include figures for DSWE1 due to the size of the match datasets, but
provide the OLS regression coefficients for both in Table \@ref(tab:roy-coeff).

## Gardner method

Gardner method intermission handoffs are defined by the data obtained in the
overlapping period of time between two adjacent-in-time missions (details
provided in Table \@ref(tab:gardner-table)). These data are filtered to sites
that have at least one data point per year for at least 75% of the years of
overlap. The filtered data are then summarized to each mission's 1st-99th
percentile value per band, and the handoff coefficients between missions are
defined by the second-order polynomial relationship between them. When n is
high, quantiles are equivalent to describing the central tendency of the
distribution of the x:y relationship due to the Central Limit Theorem. Because
we use a second-order polynomial, all values outside of the 1st and 99th
percentile values used to define the intermission handoff are flagged
`corr7_flag`, `corr8_flag` and should be used with extreme caution. All `corr7`
and `corr8` relationships and their derived coefficients are provided in the
sections below.

One additional consideration when using the Gardner method is, even when the
number of observations are is high, if there is a difference between total scene
numbers, there may be systematic differences ingrained into the coefficients
since x and y are not truly 'random samples', impacting the validity of the
Central Limit Theorem. An example of possible systematic differences could be
fewer observations from non-CONUS locations in Landsat 5 due to data
transmission errors. We did not investigate the differences in n listed in Table
\@ref(tab:gardner-table) to determine what, if any, systematic differences are
present between the two missions. We provide the Gardner method handoffs for
continuity with the riverSR product for users who would like that
interoperability.

## Establishing Landsat 4/5 and Landsat 8/9 similarity {#establish}

## Implementing Roy Handoffs

```{r roy-table}
roy_handoff_table <- tibble("Early mission" = c("Landsat 5", "Landsat 7"),
                                "Late mission" = c("Landsat 7", "Landsat 8"),
                                "n matches" = c(c(format(nrow(d_LS57_DSWE1_matches), big.mark = ","), format(nrow(d_LS89_DSWE1_matches), big.mark = ","))))
knitr::kable(roy_handoff_table, caption = "Summary of mission data used to create the Roy method intermission handoff.")
```

```{r}
create_roy_comp <- function(data, x_col, y_col, title, x_label, y_label, binwidth) {
  ggplot(data, aes(x = .data[[x_col]], y = .data[[y_col]])) +
    geom_bin2d(aes(fill = after_stat(count)), binwidth = binwidth) + 
    scale_fill_viridis_c(name = "Density", alpha = 0.5) + 
    geom_abline(intercept = 0, slope = 1, color = "grey", lty = 2) + 
    geom_smooth(method = "lm", se = FALSE, color = "red") +
coord_fixed(ratio = 1,
      xlim = c(min(data[[x_col]], data[[y_col]]), max(data[[x_col]], data[[y_col]])),
      ylim = c(min(data[[x_col]], data[[y_col]]), max(data[[x_col]], data[[y_col]]))) +
        labs(title = title, 
         x = x_label, y = y_label) +
    theme_bw()
}
```

**Landsat 5 to 7 correction (`roy_corr7`)**

```{r roy-ls57-fig, fig.dim=c(3,3), fig.ncol=2, fig.cap = "Roy method intermission handoffs of Landsat 5 relative to Landsat 7."}
plots <- pmap(.l = list(list(d_LS57_DSWE1_matches),
               list("i.med_Red", "i.med_Green", "i.med_Blue", "i.med_Nir", 
                    "i.med_Swir1", "i.med_Swir2", "i.med_SurfaceTemp"),
               list("med_Red", "med_Green", "med_Blue", "med_Nir", 
                    "med_Swir1", "med_Swir2", "med_SurfaceTemp"), 
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 5"),
               list("Landsat 7"),
               list(c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.1, 0.1))),
     .f = create_roy_comp
     )
plots 
```

**Landsat 8 to 7 correction (`roy_corr7`)**

```{r roy-ls87-fig, fig.dim=c(3,3), fig.ncol=2, fig.cap = "Roy method intermission handoffs of Landsat 8 relative to Landsat 7."}
plots <- pmap(.l = list(list(d_LS78_DSWE1_matches),
               list("med_Red", "med_Green", "med_Blue", "med_Nir", 
                    "med_Swir1", "med_Swir2", "med_SurfaceTemp"), 
               list("i.med_Red", "i.med_Green", "i.med_Blue", "i.med_Nir", 
                    "i.med_Swir1", "i.med_Swir2", "i.med_SurfaceTemp"),
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 7"),
               list("Landsat 8"),
               list(c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.1, 0.1))),
     .f = create_roy_comp
     )
plots 
```

**Landsat 7 to 8 correction (`roy_corr8`)**

```{r roy-ls78-fig, fig.dim=c(3,3), fig.ncol=2, fig.cap = "Roy method intermission handoffs of Landsat 7 relative to Landsat 8."}
plots <- pmap(.l = list(list(d_LS78_DSWE1_matches),
               list("i.med_Red", "i.med_Green", "i.med_Blue", "i.med_Nir", 
                    "i.med_Swir1", "i.med_Swir2", "i.med_SurfaceTemp"),
               list("med_Red", "med_Green", "med_Blue", "med_Nir", 
                    "med_Swir1", "med_Swir2", "med_SurfaceTemp"), 
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 8"),
               list("Landsat 7"),
               list(c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.1, 0.1))),
     .f = create_roy_comp
     )
plots 
```

**Landsat 9 to 8 correction (`roy_corr8`)**

```{r roy-ls98-fig, fig.dim=c(3,3), fig.ncol=2, fig.cap = "Roy method intermission handoffs of Landsat 9 relative to Landsat 8."}
plots <- pmap(.l = list(list(d_LS89_DSWE1_matches),
               list("med_Red", "med_Green", "med_Blue", "med_Nir", 
                    "med_Swir1", "med_Swir2", "med_SurfaceTemp"), 
               list("i.med_Red", "i.med_Green", "i.med_Blue", "i.med_Nir", 
                    "i.med_Swir1", "i.med_Swir2", "i.med_SurfaceTemp"),
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 9"),
               list("Landsat 8"),
               list(c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.001, 0.001), c(0.001, 0.001),
                    c(0.1, 0.1))),
     .f = create_roy_comp
     )
plots 
```

## Implementing Gardner Handoffs

```{r gardner-table}
gardner_handoff_table <- tibble("Early mission" = c("Landsat 5", "Landsat 7", "Landsat 8"),
                                "Late mission" = c("Landsat 7", "Landsat 8", "Landsat 9"),
                                "Correction type" = c("corr7", "corr7, corr8", "corr8"),
                                "Overlap Start" = c("1999-04-15", "2013-02-11", "2021-09-27"),
                                "Overlap End" = c("2013-02-11", "2022-04-16", "2024-12-31"),
                                "n Scenes of Early Mission" = c(format(d_LS5_forLS57corr_quantiles$n_scenes[1], big.mark = ","), 
                                                                format(d_LS7_forLS78corr_quantiles$n_scenes[1], big.mark = ","),
                                                                format(d_LS8_forLS89corr_quantiles$n_scenes[1], big.mark = ",")),
                                "n Scenes of Late Mission" = c(format(d_LS7_forLS57corr_quantiles$n_scenes[1], big.mark = ","), 
                                                               format(d_LS8_forLS78corr_quantiles$n_scenes[1], big.mark = ","), 
                                                               format(d_LS9_forLS89corr_quantiles$n_scenes[1], big.mark = ",")))

knitr::kable(gardner_handoff_table, caption = "Summary of missions for handoffs calculated by the Gardner method. Note that *n* is summarized by scene, not by number of individual site observations contributing to the quantiles.")
```

```{r}
plot_gard_quants <- function(data, x_col, y_col, title, x_label, y_label) {
  ggplot(data, aes(x = .data[[x_col]], y = .data[[y_col]], color = dswe)) +
    geom_point() +
    facet_grid(. ~ dswe) +
    geom_abline(intercept = 0, slope = 1, color = "grey", lty = 2) + 
    geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = FALSE, alpha = 0.5) +
    coord_fixed(ratio = 1,
      xlim = c(min(data[[x_col]], data[[y_col]]), max(data[[x_col]], data[[y_col]])),
      ylim = c(min(data[[x_col]], data[[y_col]]), max(data[[x_col]], data[[y_col]]))) +
    labs(title = title, 
         x = x_label, y = y_label) +
    theme_bw() +
    scale_color_viridis_d()
}
```

**Landsat 5 to 7 correction (`gard_corr7`)**

```{r gard-ls57-fig, fig.width=6}
d_LS5_forLS57corr_quantiles <- d_LS5_forLS57corr_quantiles %>% 
  select(quantile, med_Red:med_SurfaceTemp, dswe) %>% 
  rename_with(~ paste0('LS5_', .), .cols = med_Red:med_SurfaceTemp)
d_LS7_forLS57corr_quantiles <- d_LS7_forLS57corr_quantiles %>% 
  select(quantile, med_Red:med_SurfaceTemp, dswe) %>% 
  rename_with(~ paste0('LS7_', .), .cols = med_Red:med_SurfaceTemp)
LS57 <- full_join(d_LS5_forLS57corr_quantiles, d_LS7_forLS57corr_quantiles)
plots <- pmap(.l = list(list(LS57),
               list("LS5_med_Red", "LS5_med_Green", "LS5_med_Blue", "LS5_med_Nir", 
                    "LS5_med_Swir1", "LS5_med_Swir2", "LS5_med_SurfaceTemp"),
               list("LS7_med_Red", "LS7_med_Green", "LS7_med_Blue", "LS7_med_Nir", 
                    "LS7_med_Swir1", "LS7_med_Swir2", "LS7_med_SurfaceTemp"), 
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 5"),
               list("Landsat 7")),
     .f = plot_gard_quants
     )
plots
```

**Landsat 8 to 7 correction (`gard_corr7`)**

```{r gard-ls87-fig, fig.width=6}
d_LS7_forLS78corr_quantiles <- d_LS7_forLS78corr_quantiles %>% 
  select(quantile, med_Red:med_SurfaceTemp, dswe) %>% 
  rename_with(~ paste0('LS7_', .), .cols = med_Red:med_SurfaceTemp)
d_LS8_forLS78corr_quantiles <- d_LS8_forLS78corr_quantiles %>% 
  select(quantile, med_Red:med_SurfaceTemp, dswe) %>% 
  rename_with(~ paste0('LS8_', .), .cols = med_Red:med_SurfaceTemp)
LS78 <- full_join(d_LS7_forLS78corr_quantiles, d_LS8_forLS78corr_quantiles)
plots <- pmap(.l = list(list(LS78),
               list("LS8_med_Red", "LS8_med_Green", "LS8_med_Blue", "LS8_med_Nir", 
                    "LS8_med_Swir1", "LS8_med_Swir2", "LS8_med_SurfaceTemp"),
               list("LS7_med_Red", "LS7_med_Green", "LS7_med_Blue", "LS7_med_Nir", 
                    "LS7_med_Swir1", "LS7_med_Swir2", "LS7_med_SurfaceTemp"), 
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 8"),
               list("Landsat 7")),
     .f = plot_gard_quants
     )
plots
```

**Landsat 7 to 8 correction** **(`gard_corr8`)**

```{r gard-ls78-fig, fig.width=6}
plots <- pmap(.l = list(list(LS78),
               list("LS7_med_Red", "LS7_med_Green", "LS7_med_Blue", "LS7_med_Nir", 
                    "LS7_med_Swir1", "LS7_med_Swir2", "LS7_med_SurfaceTemp"), 
               list("LS8_med_Red", "LS8_med_Green", "LS8_med_Blue", "LS8_med_Nir", 
                    "LS8_med_Swir1", "LS8_med_Swir2", "LS8_med_SurfaceTemp"),
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 7"),
               list("Landsat 8")),
     .f = plot_gard_quants
     )
plots 
```

**Landsat 9 to 8 correction (`gard_corr8`)**

```{r gard-ls98-fig, fig.width=6}
d_LS8_forLS89corr_quantiles <- d_LS8_forLS89corr_quantiles %>% 
  select(quantile, med_Red:med_SurfaceTemp, dswe) %>% 
  rename_with(~ paste0('LS8_', .), .cols = med_Red:med_SurfaceTemp)
d_LS9_forLS89corr_quantiles <- d_LS9_forLS89corr_quantiles %>% 
  select(quantile, med_Red:med_SurfaceTemp, dswe) %>% 
  rename_with(~ paste0('LS9_', .), .cols = med_Red:med_SurfaceTemp)
LS89 <- full_join(d_LS8_forLS89corr_quantiles, d_LS9_forLS89corr_quantiles)

plots <- pmap(.l = list(list(LS89),
               list("LS9_med_Red", "LS9_med_Green", "LS9_med_Blue", "LS9_med_Nir", 
                    "LS9_med_Swir1", "LS9_med_Swir2", "LS9_med_SurfaceTemp"), 
               list("LS8_med_Red", "LS8_med_Green", "LS8_med_Blue", "LS8_med_Nir", 
                    "LS8_med_Swir1", "LS8_med_Swir2", "LS8_med_SurfaceTemp"),
               list("Red", "Green", "Blue", "NIR", "SWIR 1", "SWIR 2", "Surface Temperature"),
               list("Landsat 9"),
               list("Landsat 8")),
     .f = plot_gard_quants
     )
plots
```
